Differences in Homicide Case Characteristics Indicates Why Justice is Not Served

### 2024-12-02 paper/paper.qmd
Removed Data Evaluation in Appendix
Replace everything regarding New York to Chicago in paper

---
title: "Differences in Homicide Case Characteristics Indicates Why Justice is Not Served"
subtitle: "An analysis of solved and unsolved homicides from 2010 to 2017 in the United States's 2 largest cities, New York and Los Angeles"
author: 
  - Emily Su
thanks: "Code and data are available at: [https://github.com/moonsdust/unsolved-murders](https://github.com/moonsdust/unsolved-murders)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

@@

This led to us investigate the following question in our paper: what are the differences in homicide case information like the year and city the homicide took place and the victims’ perceived characteristics (age, sex, and race) between solved and unsolved homicides in 2 of the largest cities in the United States, New York and Los Angeles, from 2010 to 2017?

@@

The raw dataset we retrieved from The Washington Post using a script that downloads the CSV file from their GitHub repository contains 52179 observations with one observation being a homicide case. However, since we narrowed our scope to focus on the two of the most populated US cities, Los Angeles and New York, the number of observations in our cleaned dataset ended up being 2820. Our data look as follows:

@@

However, our dataset has limitations. There was only data available from 2010 onwards for both New York and Los Angeles provided by The Washington Post and so that limited the number of homicides we could look at. Also, not all victims were able to be identified in some homicide cases and unknown attributes of the victim such as their age, sex, and gender were indicated with the text "Unknown" in the dataset. However, this causes issues with data type compatibility such as the value "Unknown" being a character type being under the victim_age column, which has a data type of integer. This would lead to issues with our model providing accurate estimates. We decided to remove cases during our data cleaning where victim's demographic information is missing at least one of the three columns, victim_age, victim_sex, victim_race.

@@

- How homicide cases are recorded by the New York Police and Los Angeles Police 
- How The Washington Post obtain the data and what did they do

@@

We define $y_i$ to be the status of the homicide case where 1 means the homicide is unsolved (case is still open / been closed without arrest) and 0 means the homicide is solved (case has been closed with arrest). $\pi_i$ represents the probability of the homicide being solved. $\mbox{logit}(\pi_i)$ indicates the log-odds of the homicide being unsolved. Now looking at the coefficients $b_i$ and predictor variables, $\beta_0$ is the intercept of our model and is the log-odds when all predictor variables are equal to 0. $\mbox{victim\_race}_i$ signifies the race of the victim, which could be either "Asian", "Black", "Hispanic", "White", and "Other". In our model, we use "White" as the baseline for $\mbox{victim\_race}_i$ to see if being part of a minority impacts if the case goes unsolved or not. $\beta_1$ is the coefficient that represents the log-odds when $\mbox{victim\_race}_i$ changes. $\mbox{victim\_age}_i$ is the victim's age, which is a whole number. $\beta_2$ is the log-odds when $\mbox{victim\_age}_i$ increases by 1 year. $\mbox{victim\_sex}_i$ represents the sex of the victim, where in the dataset it is either "female" or "male" and $\beta_3$ is the log-odds when $\mbox{victim\_sex}_i$ changes. $\mbox{city}_i$ is the city the victim was reported to be killed in, which from our dataset could be either "New York" or "Los Angeles". $\beta_4$ is the coefficient that stands for the log-odds when $\mbox{city}_i$ changes. We define $\mbox{year}_i$ to be the year the homicide was reported to have occured from 2010 to 2017. $\beta_5$ indicates the log-odds when $\mbox{year}_i$ increases by 1 year.

@@

Performing root mean square deviation (RSME) calculations on our model in @tbl-model-result yields a RSME value of about 0.49. RSME represents the difference between the model's predicted values and observed values with 0 meaning that the predicted and observed values are the same [@citejim]. Since the model's RSME value is close to 0, we can say our model is able to predict values with less error compared to other models that have a higher RSME value.

@@

## Differences in Homicide Case Information Between Solved and Unsolved Cases in New York and Los Angeles (2010 to 2017)

### Date (Month and Year)
```{r}
#| label: fig-month-bar
#| fig-cap: "Number of solved and unsolved homicides across the 12 months of a year in Los Angeles and New York (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-year-bar
#| fig-cap: "Number of solved and unsolved homicides from 2010 to 2017 in Los Angeles and New York"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-year-month-heatmap
#| fig-cap: "Number of solved and unsolved homicides from January to December from 2010 to 2017 in Los Angeles and New York"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-city-bar
#| fig-cap: "Proportion of solved and unsolved homicides in Los Angeles and New York (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-city-table
#| tbl-cap: "Proportion and number of solved and unsolved homicides in Los Angeles and New York (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Disposition
```{r}
#| label: fig-disposition-bar
#| fig-cap: "Disposition of homicide cases in New York and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-disposition-table
#| tbl-cap: "Disposition of homicide cases in New York and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Victim's Age 
```{r}
#| label: fig-victim-age-histogram
#| fig-cap: "Distribution of victim's age in solved and unsolved homicides in New York and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Victim's Sex
```{r}
#| label: fig-victim-sex-bar
#| fig-cap: "Proportion of homicide cases per sex in New York and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-victim-sex-table
#| tbl-cap: "Proportion and number of homicide cases per sex in New York and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Victim's Race 
```{r}
#| label: fig-victim-race-bar
#| fig-cap: "Number of homicide cases per race in New York and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-victim-race-table
#| tbl-cap: "Number of homicide cases per sex in New York and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-month-table
#| tbl-cap: "Number of solved and unsolved homicides across the 12 months of a year in Los Angeles and New York (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-year-table
#| tbl-cap: "Number of solved and unsolved homicides from 2010 to 2017 in Los Angeles and New York"
#| echo: false
#| warning: false
#| message: false

@@

\newpage
## Idealized Survey and Methodology {#sec-ideal-survey}
- Link to literature

### Idealized Survey Objectives

### Sampling Approach

@@

### Idealized Survey Design

@@

### Limitations

@@

\newpage 
## Overview and Evaluation of The Washington Post’s Dataset {#sec-data-evalulation}
- TODO: Make sure to link evaluation to literature

### Overview

### Strengths and limitations
- what is good and bad about the sampling.

>>>>

---
title: "Differences in Homicide Case Characteristics Indicates Why Justice is Not Served"
subtitle: "An analysis of solved and unsolved homicides from 2010 to 2017 in one of the United States's 2 largest cities, Chicago and Los Angeles"
author: 
  - Emily Su
thanks: "Code and data are available at: [https://github.com/moonsdust/unsolved-murders](https://github.com/moonsdust/unsolved-murders)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

@@

This led to us investigate the following question in our paper: what are the differences in homicide case information like the year and city the homicide took place and the victims’ perceived characteristics (age, sex, and race) between solved and unsolved homicides in one of the 2 of the largest cities in the United States, Chicago and Los Angeles, from 2010 to 2017?

@@

The raw dataset we retrieved from The Washington Post using a script that downloads the CSV file from their GitHub repository contains 52179 observations with one observation being a homicide case. However, since we narrowed our scope to focus on the two of the most populated US cities, Los Angeles and Chicago, the number of observations in our cleaned dataset ended up being 2820. Our data look as follows:

@@

However, our dataset has limitations. There was only data available from 2010 onwards for Los Angeles provided by The Washington Post and so that limited the number of homicides we could look at for both Chicago and Los Angeles. Also, not all victims were able to be identified in some homicide cases and unknown attributes of the victim such as their age, sex, and gender were indicated with the text "Unknown" in the dataset. However, this causes issues with data type compatibility such as the value "Unknown" being a character type being under the victim_age column, which has a data type of integer. This would lead to issues with our model providing accurate estimates. We decided to remove cases during our data cleaning where victim's demographic information is missing at least one of the three columns, victim_age, victim_sex, victim_race.

@@

- How homicide cases are recorded by the Chicago Police and Los Angeles Police 
- How The Washington Post obtain the data and what did they do

The Federal Bureau of Investigation (FBI) has a program called the Uniform Crime Reporting (UCR) Program to generate statistics for the public [@citeFBIinfo]. The FBI originally had a system under the UCR called Summary Reporting System (SRS), which obtained details about different crimes taking place from law enforcement agencies nationwide such as victim information [@citeFBIinfo]. However, the SRS was replaced with a new system called the National Incident-Based Reporting System (NIBR) in 2021 that obtained more details about various crimes [@citeFBIinfo]. For The Los Angeles Police Department, after a homicide case occurs, the investigating team handwrites information into physical crime reports with details like the type of crime, the premise the crime occured at, and the age and ethnicity of the victim [@citeLApolice]. The crime reports are then transcribed into a digital format and then sent to the SRS monthly [@citeLApolice].

@@

We define $y_i$ to be the status of the homicide case where 1 means the homicide is unsolved (case is still open / been closed without arrest) and 0 means the homicide is solved (case has been closed with arrest). $\pi_i$ represents the probability of the homicide being solved. $\mbox{logit}(\pi_i)$ indicates the log-odds of the homicide being unsolved. Now looking at the coefficients $b_i$ and predictor variables, $\beta_0$ is the intercept of our model and is the log-odds when all predictor variables are equal to 0. $\mbox{victim\_race}_i$ signifies the race of the victim, which could be either "Asian", "Black", "Hispanic", "White", and "Other". In our model, we use "White" as the baseline for $\mbox{victim\_race}_i$ to see if being part of a minority impacts if the case goes unsolved or not. $\beta_1$ is the coefficient that represents the log-odds when $\mbox{victim\_race}_i$ changes. $\mbox{victim\_age}_i$ is the victim's age, which is a whole number. $\beta_2$ is the log-odds when $\mbox{victim\_age}_i$ increases by 1 year. $\mbox{victim\_sex}_i$ represents the sex of the victim, where in the dataset it is either "female" or "male" and $\beta_3$ is the log-odds when $\mbox{victim\_sex}_i$ changes. $\mbox{city}_i$ is the city the victim was reported to be killed in, which from our dataset could be either "Chicago" or "Los Angeles". $\beta_4$ is the coefficient that stands for the log-odds when $\mbox{city}_i$ changes. We define $\mbox{year}_i$ to be the year the homicide was reported to have occured from 2010 to 2017. $\beta_5$ indicates the log-odds when $\mbox{year}_i$ increases by 1 year.

@@

Performing root mean square deviation (RSME) calculations on our model in @tbl-model-result yields a RSME value of about 0.45. RSME represents the difference between the model's predicted values and observed values with 0 meaning that the predicted and observed values are the same [@citejim]. Since the model's RSME value is close to 0, we can say our model is able to predict values with less error compared to other models that have a higher RSME value.

@@

## Differences in Homicide Case Information Between Solved and Unsolved Cases in Chicago and Los Angeles (2010 to 2017)

### Date (Month and Year)
```{r}
#| label: fig-month-bar
#| fig-cap: "Number of solved and unsolved homicides across the 12 months of a year in Los Angeles and Chicago (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-year-bar
#| fig-cap: "Number of solved and unsolved homicides from 2010 to 2017 in Los Angeles and Chicago"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-year-month-heatmap
#| fig-cap: "Number of solved and unsolved homicides from January to December from 2010 to 2017 in Los Angeles and Chicago"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-city-bar
#| fig-cap: "Proportion of solved and unsolved homicides in Los Angeles and Chicago (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-city-table
#| tbl-cap: "Proportion and number of solved and unsolved homicides in Los Angeles and Chicago (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Disposition
```{r}
#| label: fig-disposition-bar
#| fig-cap: "Disposition of homicide cases in Chicago and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-disposition-table
#| tbl-cap: "Disposition of homicide cases in Chicago and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Victim's Age 
```{r}
#| label: fig-victim-age-histogram
#| fig-cap: "Distribution of victim's age in solved and unsolved homicides in Chicago and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Victim's Sex
```{r}
#| label: fig-victim-sex-bar
#| fig-cap: "Proportion of homicide cases per sex in Chicago and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-victim-sex-table
#| tbl-cap: "Proportion and number of homicide cases per sex in Chicago and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

### Victim's Race 
```{r}
#| label: fig-victim-race-bar
#| fig-cap: "Number of homicide cases per race in Chicago and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-victim-race-table
#| tbl-cap: "Number of homicide cases per sex in Chicago and Los Angeles (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-month-table
#| tbl-cap: "Number of solved and unsolved homicides across the 12 months of a year in Los Angeles and Chicago (2010 to 2017)"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-year-table
#| tbl-cap: "Number of solved and unsolved homicides from 2010 to 2017 in Los Angeles and Chicago"
#| echo: false
#| warning: false
#| message: false

@@

\newpage
## Idealized Survey and Methodology {#sec-ideal-survey}

### Idealized Survey Objectives
The objective of our survey is to collect information from Chicago Police Department and Los Angeles Police Department about their education background, reason for closing a homicide case without arrest, difficult homicide case they have done, how many cases they have to cover at a time, etc.

regarding resources available

The sampling approach we plan to take is stratified sampling, which is a type of probabilistic sampling [@citetellingstories]

@@

### Idealized Survey Design
- Using an online survey in cases.

@@

### Limitations
- what is good and bad about the sampling.

@@
